---
title: "fitgen"
author: ""  
date: "`r Sys.Date()`"  
address: Departament of Informatics and Departament of Statistics and Operation Research  
format: 
    revealjs:  
      scrollable: true
      echo: true
      embed-resources: true
    pdf:
       toc: true
       number-sections: false
       colorlinks: true  
---

# Packages and data

The following packages are required to run the examples in this vignette.

## Packages  

```{r}
#| label: fg1
pacman::p_load(fitgen,SummarizedExperiment,ggplot2,edgeR,TCGAbiolinks)

```

## Example dataset

The functions in `fitgen` require a response matrix containing feature-level
counts and an associated model matrix containing the sample-level covariates.

RNA-seq data are commonly stored in a `RangedSummarizedExperiment` object,
provided by the `SummarizedExperiment` package. In this vignette, we use a
reduced version of the TCGA COAD dataset (TCGA-COAD).
The complete dataset contains gene-level expression counts together with
clinical and molecular information for tumour and normal breast tissue
samples.

The code used to download and prepare the original TCGA-COAD dataset is shown
below for reproducibility. Because downloading and processing the complete
dataset is computationally intensive, this code is not evaluated when the
vignette is rendered.
```{r}
#| label: fg2
#| eval: false  
options(timeout = 600)  
query <- GDCquery(
  project = "TCGA-COAD",
  data.category = "Transcriptome Profiling",
  data.type = "Gene Expression Quantification",
  workflow.type = "STAR - Counts"
)
GDCdownload(
  query,
  method = "api",           
  files.per.chunk = 30, 
  directory = "GDCdata" 
)
y <- GDCprepare(query, directory = "GDCdata") ##
save(y,file="tcga_coad.rda")
```

### Filtering and normalization

Lowly expressed features are removed before model fitting. Features 
with duplicated gene symbols are also excluded so that each retained 
symbol is represented only once.
 
Differences in library composition and library size are accounted for 
by estimating normalization factors with the TMM method implemented in 
`edgeR`.
These factors are stored in the `DGEList` object; the original count 
matrix is retained. Gene identifiers are also processed to remove the 
version suffix from the Ensembl identifiers.

The resulting object is saved for use in the subsequent preparation of 
the reduced example dataset.

```{r}
#| label: fig3 
#| eval: false
load("tcga_coad.rda")
sel1 = which(colData(y)$sex_at_birth !="unknown")
sel2 = which(!is.element(colData(y)$sample_type,
                         c("Metastatic","Recurrent Tumor")))
y = y[,intersect(sel1,sel2)]
colData(y)$sex_at_birth = factor(colData(y)$sex_at_birth)
colData(y)$sample_type = factor(colData(y)$sample_type)
```
### Construction of the reduced dataset

To keep the examples computationally manageable, a reduced dataset is
constructed from the complete TCGA-COAD data.

Only the clinical variables used in the examples are retained. Samples with
missing values for these variables are excluded, as are metastatic and
recurrent tumour samples. Lowly expressed features are then filtered out.

For the construction of this reduced dataset, adjusted count values are
obtained from an `edgeR` object after estimating a common dispersion
parameter. These values are rounded and stored in the assay of the reduced
object. Finally, a random subset of 100 features is selected using a fixed
random seed so that the example is reproducible.

This preprocessing step is shown for documentation purposes and is not
evaluated when the vignette is rendered.
```{r}
#| label: fg3
#| eval: false
vars = c("sex_at_birth","sample_type")
colData(y) = colData(y)[,vars]
x0 = colData(y)[,vars]
selcol = which(complete.cases(x0))
to_keep = rowSums(cpm(assay(y)) > 0.5) > (ncol(y)/100)
table(to_keep)
y = y[to_keep,]
dge = edgeR::DGEList(counts=assay(y))
dge.c = edgeR::estimateCommonDisp(dge) 
dge.c$common.dispersion ## 1.041049
counts0 = dge.c$pseudo.counts
rownames(counts0) = rownames(assay(y))
colnames(counts0) = colnames(assay(y))
assay(y) = round(counts0)
set.seed(1234)
y = y[sample(nrow(y),100),]
save(y,file="y.rda")
```

The resulting reduced object, `y`, is included in the `fitgen` package and can be loaded 
directly as follows.

```{r}
#| label: fg4
data(y)
```
The rows of `y` correspond to features and the columns correspond to samples.
The assay contains the response values used in the examples, whereas the sample-level variables 
required to construct the model matrix are stored in `colData(y)`.

# Model fitting

## Fitting a model to a single row

We first illustrate the main fitting functions using a single row of the
response matrix. The same sample-level covariates are used in all models.

The variables `sex_at_birth` and `sample_type` are included as common predictors.
The corresponding model matrix, `xcommon`, is constructed from the sample
metadata stored in `colData(y)`.

```{r}
#| label: fg5
vars = c("sex_at_birth","sample_type")
ff = paste0("~ ",paste(vars,collapse="+"))
ff = as.formula(ff)
xcommon = model.matrix(ff,data=colData(y))
summary(as.data.frame(colData(y))[,vars])

```
We select a  row  from the assay matrix. The resulting vector, `y0`,
contains the response values for this feature across all samples.

A binary response, `y1`, is also constructed to illustrate logistic
regression. Observations above the 60th percentile of `y0` are assigned the
value 1, and the remaining observations are assigned the value 0. This
dichotomization is used only for illustration and is not intended as a
general preprocessing strategy for RNA-seq data.

```{r}  
#| label: fg6
i = 50
yt = assay(y)
y0 = yt[i,]
quantile(y0,probs=.6)
y1 = (y0 > quantile(y0,probs=.6))* 1. ## For logistic regression
x0 =  colData(y)[,vars]
df0 = data.frame(x0,y0,y1)
names(df0)

```

## Poisson model

We first fit a Poisson generalized linear model using `stats::glm()`. This
fit serves as a reference for comparison with the corresponding functions
implemented in `fitgen`.

The model uses a log link and assumes that the conditional variance is equal
to the conditional mean:
$$
\operatorname{Var}(Y_j \mid \mathbf{x}_j) = \mu_j.
$$
The coefficients, deviance, and fitted values obtained from this reference
model are displayed below.

```{r}
#| label: fg7
fit0 = glm(y0~ sex_at_birth + sample_type,family = "poisson",data=df0)
coef(fit0)
deviance(fit0)
head(fitted(fit0))

```

We next fit the same Poisson model using `fitgen::StGLM()`. The response
vector, model matrix, distribution, and offset are supplied explicitly.
Setting `retmu = TRUE` indicates that the fitted means should be returned.

The resulting object contains the model deviance, the estimated regression
coefficients, and the fitted means. Because the same response, model matrix,
link function, and variance specification are used, the results can be
compared directly with those obtained from `stats::glm()`.

```{r}
#| label: fg8
Fit0 = StGLM(y=y0,X=xcommon,model="poisson",offset=rep(0,length(y0)),
             warn_me=TRUE,retmu=TRUE)
Fit0$Dev ## deviance
Fit0$beta ## coefficients
head(Fit0$muest) ## fitted values (estimates for the observations and means)

```
The function `fitgen::glmRow()` provides an interface similar to that of
`stats::glm()`. Standard extractor functions such as `coef()`,
`coefficients()`, `deviance()`, and `fitted()` can therefore be used with
the fitted object.

Setting `is_fitted_values = TRUE` indicates that the fitted values should be
stored and returned.

```{r}
#| label: fg9
fit = glmRow(x=xcommon,y=y0,family="poisson",is_fitted_values=TRUE)
coef(fit)
coefficients(fit)
deviance(fit)
head(fitted(fit))

```
The three procedures fit the same Poisson model to the selected row. Their
estimated coefficients, deviances, and fitted values should therefore agree,
up to minor numerical differences arising from the fitting algorithms.

## Negative binomial model

We next fit an NB2 regression model to the same response vector. Under the
NB2 parameterization used here, the conditional variance is
$$
\operatorname{Var}(Y_j \mid \mathbf{x}_j)
  = \mu_j + \phi \mu_j^2,
$$
where $\phi$ is the negative binomial dispersion parameter.

We first fit the model using `stats::glm()` with the negative binomial family
provided by the `MASS` package. This fit serves as a reference for comparison
with the corresponding functions implemented in `fitgen`.

The `MASS::negative.binomial()` family expresses the dispersion in terms of
`theta`. Under the parameterization used here,
$$
\theta = \frac{1}{\phi}.
$$
In this example, $\phi$ is fixed at $0.7983342$, so the corresponding value
`theta = 1 / 0.7983342` is supplied to `MASS::negative.binomial()`.

The estimated coefficients, deviance, and fitted values are displayed below.
```{r}
#| label: fg10
fit0 = glm(y0~ sex_at_birth + sample_type,
           family = MASS::negative.binomial(theta=1/0.7983342),data=df0)
coef(fit0)
deviance(fit0)
head(fitted(fit0))

```

We next fit the same NB2 model using `fitgen::StGLM()`. The response vector,
model matrix, distribution, offset, and dispersion parameter are supplied
explicitly.

The argument `Phi` represents the NB2 dispersion parameter $\phi$. In this
fit, $\phi$ is treated as fixed while the regression coefficients are
estimated. Setting `retmu = TRUE` indicates that the fitted means should be
returned.

The resulting object contains the model deviance, the estimated regression
coefficients, and the fitted means. Because the same response, model matrix,
link function, and fixed value of $\phi$ are used, the results can be compared
directly with those obtained from `stats::glm()`.

```{r}
#| label: fg11
Fit0 = StGLM(y=y0,X=xcommon,model="negbin",offset=rep(0,length(y0)),
             warn_me=TRUE,retmu=TRUE,Phi=0.7983342)
Fit0$Dev ## deviance
Fit0$beta ## coefficients
head(Fit0$muest) ## fitted values (estimates for the observations and means)

```
The function `fitgen::NB2Row()` provides a single-row interface similar to
that of `stats::glm()`. Standard extractor functions such as `coef()`,
`coefficients()`, `deviance()`, and `fitted()` can therefore be used with
the fitted object.

The argument `phi` represents the fixed NB2 dispersion parameter $\phi$.

Setting `is_fitted_values = TRUE` indicates that the fitted values should be
stored and returned.

```{r}
#| label: fg12
fit = NB2Row(x=xcommon,y=y0,phi=0.7983342,is_fitted_values=TRUE)
coef(fit)
coefficients(fit)
deviance(fit)
head(fitted(fit))

```
The three procedures fit the same NB2 model to the selected row using the same
fixed value of $\phi$. Their estimated coefficients, deviances, and fitted
values should therefore agree, up to minor numerical differences arising
from the fitting algorithms.


## Binomial model: logistic regression

Finally, we illustrate the use of `fitgen::glmRow()` with a binary response.
The variable `y1` indicates whether each observation is above the 60th
percentile of the response values for the selected feature.

We first fit a logistic regression model using `stats::glm()`. We then fit
the same model using `fitgen::glmRow()`. Both fits use a binomial response
with a logit link and the same model matrix.

The estimated coefficients and deviance are extracted from both fitted
objects. For the model fitted with `glmRow()`, setting
`is_fitted_values = TRUE` indicates that the fitted probabilities should be
stored and returned.

```{r}
#| label: fg13
fit1 = glm(y1~ sex_at_birth + sample_type,family = "binomial",data=df0)
coef(fit1)
deviance(fit1)

fit2 = glmRow(x=xcommon,y=y1,family="binomial",is_fitted_values=TRUE)
coefficients(fit2)
coef(fit2)
deviance(fit2)
head(fitted(fit2))

```
The two procedures fit the same logistic regression model to the selected
row. Their estimated coefficients and deviances should therefore agree, up
to minor numerical differences arising from the fitting algorithms. The
fitted values returned by `glmRow()` are the estimated probabilities of
observing a value equal to 1.

This example is included only to illustrate the use of the row-wise interface
with a binomial response. The dichotomization of the count data should not be
interpreted as a recommended procedure for differential expression analysis,
as it discards information from the original response.

# Fitting one model per row

## Poisson models

We now fit a separate generalized linear model to each row of the response
matrix `yt`. All row-wise models include the same sample-level covariates,
represented by the common model matrix `xcommon`.

Two model specifications are considered. The first includes only the
predictors in `xcommon`. The second extends `xcommon` with row-specific
predictors derived from neighboring features. For each target row, the
expression profiles of the neighbors identified by
`fitgen::find_neighbours()` are added as additional columns to `xcommon`.
The resulting extended model matrix is therefore specific to each target row.

The model containing only `xcommon` describes the association between the
sample-level covariates and the response for each feature without conditioning
on neighboring features. By contrast, the neighborhood-adjusted model
describes this association conditional on the expression profiles of the
selected neighbors.

Comparing the two specifications makes it possible to assess whether
conditioning on locally related features changes the fitted means, residual
variability, or estimated effects of the sample-level covariates. The two
specifications should therefore be interpreted as related but distinct
models.

In this example, five neighboring features are used to extend `xcommon` for
each row-wise model. Because identifying the neighbors can be computationally
intensive, this step is not evaluated when the vignette is rendered. The
resulting object is saved and reused in the subsequent examples.
```{r}
#| label: fg14
neighbours = find_neighbours(yt,neighbours_n=5) ## Slow processing

```
The neighbors identified for row `r i` are shown below.

```{r}
#| label: fg16
neighbours[i,]
                            
```
We now fit a separate Poisson generalized linear model to each row of `yt` 
using `fitgen::glmRows()`.

All row-wise models include the predictors in `xcommon`. For each target row, the expression 
profiles of its previously identified neighbors are added asnadditional columns to the common
model matrix. Thus, each full model contains a common component, shared by all rows, and a 
row-specific component defined by the selected neighbors.

The argument `selected` is a logical vector indicating the rows for which a model should be
fitted. In this example, all rows are selected. Setting `is_fitted_values = TRUE` indicates 
that the fitted values should be storedand returned.

Because the neighbors have already been identified, they are supplied directly through the 
`neighbours` argument. Fitting one model per row with additional neighbor-based predictors 
can be computationally intensive.
Therefore, this step is not evaluated when the vignette is rendered, and the previously 
fitted object is saved for use in the following sections.
```{r}
#| label: fg17
fits = glmRows(x=xcommon,y=yt,family="poisson",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yt)), 
                 is_fitted_values=TRUE, ## The fitted values are  returned
                 neighbours = neighbours)

```

The object `fits` contains the row-wise Poisson fits obtained using the common
predictors and the additional neighbor-based predictors. This object belongs
to the `fitgen_Lglm` class.
```{r}
#| label: fg20
class(fits)

```
S3 methods are available for objects of class `fitgen_Lglm`. These methods
extend the usual extractor functions for generalized linear models to an
object containing multiple row-wise fits.

For each row, results can be extracted from two model specifications. With
`type = "common"`, the results correspond to the model containing only the
predictors in `xcommon`. With `type = "full"`, the results correspond to the
extended model containing both the common predictors and the row-specific
neighbor-based predictors.

The following code extracts the coefficients, fitted values, and deviances
from both model specifications. It also computes a feature-wise
quasi-dispersion estimate using the Wedderburn method.

```{r}
#| label: fg21
a = coef(fits,type="common")
a = coefficients(fits,type="common") ## Only common predictors are used
a[[1]] ## the first vector of coefficients

a = coef(fits,type="full")
a = coefficients(fits,type="full") ## Common and neighbor-based predictors are used 
a[[1]] ## the first vector of coefficients


b = fitted(fits,type="common") ## Fitted values with common predictors
head(b[[1]])  

b = fitted(fits,type="full") ## ## Fitted values with common and neighbor-based predictors
head(b[[1]]) 

c = deviance(fits,type="common") ## Deviances with common predictors
c[[1]] ## the first one
summary(unlist(c))

c = deviance(fits,type="full") ## Deviances with all predictors
c[[1]] ## the first one
summary(unlist(c))

## Quasi dispersion with common predictors  where type can be
## "Wedderburn","Farrington" or "Fletcher"
d = quasidispersion(fits,type="Wedderburn") 
d[[1]] ## the first one
summary(unlist(d))

``` 

We can also fit the row-wise Poisson models without neighbor-based predictors.
In this case, each model contains only the predictors in the common model
matrix `xcommon`.

This specification provides the unadjusted model for each feature and can be
used as a reference when assessing the effect of including neighboring
features. Because no neighbor-based predictors are included, the common and
full model specifications coincide.

As before, all rows are selected and the fitted values are requested. This
computationally intensive step is not evaluated when the vignette is rendered,
and the resulting object is saved for subsequent use.

```{r}
#| label: fg22
#| eval: false

fits = glmRows(x=xcommon,y=yt,family="poisson",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yt)), 
                 is_fitted_values=TRUE,
                 neighbours=NULL,
                 neighbours_n = 0)

```

## Logistic regression

To illustrate row-wise logistic regression, the count matrix `yt` is
dichotomized to obtain a binary response matrix `yl`. In this example, counts
greater than or equal to 100 are assigned the value 1, whereas counts below
100 are assigned the value 0.

This transformation is introduced solely to demonstrate the use of
`fitgen::glmRows()` with a binomial response. It should not be interpreted as
a generally recommended transformation for RNA-seq differential expression
analysis, since dichotomization discards information contained in the
original count data.

```{r}
#| label: fg23
yl = (yt >= 100)*1
summary(apply(yl,1,sum))

```
We now fit a separate logistic regression model to each row of the binary
response matrix `yl`.

All models include the predictors in `xcommon`. For each target row, the
expression profiles of its previously identified neighbors are added as
additional columns to the common model matrix. Thus, each full model contains
a common component and a row-specific component defined by the selected
neighbors.

The previously computed neighbors are supplied through the `neighbours`
argument. All rows are selected, and setting `is_fitted_values = TRUE`
indicates that the fitted probabilities should be stored and returned.

Because fitting one neighborhood-adjusted model per row can be computationally
intensive, this step is not evaluated when the vignette is rendered. The
resulting object is saved for subsequent inspection.

```{r}
#| label: fg24
#| eval: false

fits = glmRows(x=xcommon,y=yl,family="binomial",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yl)), 
                 is_fitted_values=TRUE, ## The fitted values are returned
                 neighbours = neighbours) ## Previously calculated

save(fits,file="~/Dropbox/fitgen-data/fits_with_neighbours_logistic_tcga_coad.rda")

```

We now load the previously fitted object containing the row-wise logistic
regression models.

```{r}
#| label: fg25
load("~/Dropbox/fitgen-data/fits_with_neighbours_logistic_tcga_coad.rda")

```

The availability of the deviance provides a simple diagnostic of whether a model produced a 
valid result for each row. The following code extracts the
deviances from the full models, which include both the common predictors and
the row-specific neighbor-based predictors.

We then count the number of missing deviance values and summarize the distribution of the
available values.


```{r}
#| label: fg26
a = deviance(fits,type="full")
table(is.na(unlist(a)))
summary(unlist(a))
```

The remaining extractor methods can be used in the same way to obtain
coefficients, fitted probabilities, and quasi-dispersion estimates from the
row-wise logistic regression models.
 

Next, we fit the same row-wise logistic regression models **without**
neighbor-based predictors. In this case, each model contains only the
predictors in the common model matrix `xcommon`.

This specification provides the unadjusted model for each feature and can be
used as a reference for assessing the effect of including neighboring
features.

```{r}
#| label: fg27
fits = glmRows(x=xcommon,y=yl,family="binomial",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yl)), 
                 is_fitted_values=TRUE, 
                 neighbours = NULL,  ## No neighbours are provided
                 neighbours_n = 0) ## Number of neighbours is null

```


## Negative binomial models

We now fit a separate NB2 regression model to each row of the response matrix
`yt`.

Under the NB2 parameterization used here, the conditional variance for each feature is

$$
\operatorname{Var}(Y_j \mid \mathbf{x}_j)
  = \mu_j + \phi \mu_j^2,
$$

where $\phi$ is the negative binomial dispersion parameter.

In this example, a feature-specific value of $\phi$ is estimated using the
tagwise dispersion method implemented in `edgeR`. This approach allows the
NB2 dispersion parameter to vary across features while borrowing information
across genes during estimation.

The dispersion estimates are obtained before fitting the row-wise NB2 models.
Because this step can be computationally intensive, it is not evaluated when
the vignette is rendered. The resulting vector of feature-specific dispersion
estimates is saved for use in the subsequent models.

```{r}
#| label: fg28
phi = NB2_phi(y=yt,method="tagwise_edger",selected=rep(TRUE,nrow(yt)))

```
We now use the previously computed tagwise dispersion estimates to fit one
NB2 model to each row of `yt`.

All models include the predictors in `xcommon`. For each target row, the
expression profiles of its previously identified neighbors are added as
additional columns to the common model matrix. The resulting full model
therefore contains both the common predictors and the row-specific
neighbor-based predictors.

The object `phi` contains the feature-specific NB2 dispersion estimates
previously obtained using the tagwise method implemented in `edgeR`. These
externally estimated values are supplied to `fitgen::NB2Rows()` and used
during the row-wise model fitting. Because `phi` is provided explicitly, the
dispersion parameters are not re-estimated at this stage.

All rows are selected, and setting `is_fitted_values = TRUE` indicates that
the fitted values should be stored and returned. Because fitting one
neighborhood-adjusted NB2 model per row can be computationally intensive,
this step is not evaluated when the vignette is rendered. The resulting
object is saved for subsequent inspection.

```{r}
#| label: fg30
#| results: false
fits = NB2Rows(x=xcommon,
                 y=yt,
                 phi=phi,
                 phi_method = "single",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yt)),  ## All rows are selected  
                 is_fitted_values=TRUE, ## The fitted values are returned
                 neighbours = neighbours)

```

The object `fits` contains the row-wise NB2 fits obtained using the
feature-specific tagwise dispersion estimates.

For each row, results can be extracted from two model specifications. With
`type = "common"`, the results correspond to the model containing only the
predictors in `xcommon`. With `type = "full"`, the results correspond to the
extended model containing both the common predictors and the row-specific
neighbor-based predictors.

The following code extracts the coefficients, fitted values, and deviances
from both model specifications. It also computes feature-wise
quasi-dispersion estimates using the Wedderburn method and NB2 dispersion
estimates using the Bliss method.

The NB2 dispersion parameter and the quasi-dispersion parameter have different
roles. The NB2 dispersion parameter $\phi$ defines the baseline quadratic
mean--variance relationship, whereas the quasi-dispersion parameter measures
the remaining variability relative to that variance specification after the
mean model has been fitted.

```{r}
#| label: fg32
a = coef(fits,type="common")
a = coefficients(fits,type="common") ## Only common predictors are used
a[[1]] ## the first vector of coefficients

a = coef(fits,type="full")
a = coefficients(fits,type="full") ## Common and neighbor-based predictors are used
a[[1]] ## the first vector of coefficients


b = fitted(fits,type="common") ## Fitted values with common predictors
head(b[[1]])  ## 

b = fitted(fits,type="full") ## Fitted values with common and neighbor-based predictors
head(b[[1]]) 

c = deviance(fits,type="common") ## Deviances with common predictors
c[[1]] ## the first one
summary(unlist(c))

c = deviance(fits,type="full") ## Deviances with all predictors
c[[1]] ## the first one
summary(unlist(c))

d = quasidispersion(fits,type="Wedderburn") ## Feature-wise quasi-dispersion estimates

d[[1]] ## the first one
summary(unlist(d))

e = dispersion(fits,type="Bliss") ## NB2 dispersion estimates using the Bliss method
e[[1]] ## ## The first dispersion estimate
summary(unlist(e))

``` 


We next fit the same row-wise NB2 models without neighbor-based predictors.
In this case, each model contains only the predictors in the common model
matrix `xcommon`.

The same feature-specific values of the NB2 dispersion parameter $\phi$ are
used as in the neighborhood-adjusted models. Therefore, comparing the two
specifications makes it possible to assess the effect of adding
neighbor-based predictors while keeping the NB2 dispersion parameter fixed.

As before, all rows are selected and the fitted values are requested. This
computationally intensive step is not evaluated when the vignette is rendered,
and the resulting object is saved for subsequent use.

```{r}
#| label: fg33
fits = NB2Rows(x = xcommon,
                 y = yt,
                 phi=phi,
                 phi_method = "single",
                 is_normalized=TRUE,
                 selected = rep(TRUE,nrow(yt)), 
                 is_fitted_values=TRUE,
                 neighbours = NULL,
                 neighbours_n = 0)

```

We now load the previously fitted object containing the row-wise NB2 models
without neighbor-based predictors.

## Fitting alternative models 
The function `fitgen::FitRows()` provides a high-level interface for fitting
alternative row-wise models within a common analytical framework.

All models use the predictors in `xcommon`. For each target row, the
expression profiles of the previously identified neighbors are added as
row-specific predictors. The externally estimated NB2 dispersion parameters
contained in `phi` are supplied to the fitting function and are not
re-estimated at this stage.

In addition to the Poisson and NB2 fits, `FitRows()` evaluates
quasi-dispersion using the method specified by `quasi_method`. In this
example, the Wedderburn estimator is used.

The arguments `alpha` and `p.adjust.method` specify the significance level
and the method used to adjust for multiple comparisons, respectively.
Setting `p.adjust.method = "BH"` applies the Benjamini--Hochberg adjustment.

Because fitting and evaluating alternative models for every row can be
computationally intensive, this step is not evaluated when the vignette is
rendered. The resulting object is saved for subsequent inspection.

```{r}
#| label: fg35
#| results: false
res = FitRows(x=xcommon,y=yt,phi  = phi,
                    is_normalized=TRUE,
                    is_fitted_values=TRUE,
                    neighbours_method="trivial",
                    neighbours = neighbours)

```

We now load the previously computed results. The returned object contains the
row-wise Poisson and NB2 fits, together with the corresponding diagnostic and
inferential summaries.

The class of the Poisson component is inspected below. We also summarize the
model classifications stored in `res$type.model` and the feature-wise
quasi-dispersion estimates obtained from the Poisson and NB2 fits.

```{r}
#| label: fg36
class(res$Poi)
summary(unlist(quasidispersion(res$Poi)))
summary(unlist(quasidispersion(res$NB2)))

```

The Poisson and NB2 quasi-dispersion estimates quantify residual variability
relative to different baseline variance functions. For the Poisson fits, the
baseline variance is the conditional mean. For the NB2 fits, it is determined
by the conditional mean and the supplied NB2 dispersion parameter.

Therefore, quasi-dispersion estimates from the Poisson and NB2 fits should be
interpreted relative to their respective baseline variance functions. Their
values are not directly comparable without taking these different variance
specifications into account.

The results returned by `FitRows()` support the comparison of feature-wise
conclusions under alternative variance and dispersion assumptions. They should
not be interpreted as identifying a universally optimal model for every
feature.
